\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Assessing economic inequality with tax data - Switzerland from 1945 to 2010}

### 2014-06-16 paper/subdocuments/data_and_methods.Rnw
minor changes methods section

@@

% Bei den Nullern erw?hnen, ab wann Angaben zur Zahl der Nuller vorliegen und ab wann nicht
% Im Internet hat es Publikationen ab 1947/48
% Es gibt auch Tabellen auf Gemeindeebene
% Die Statitstischen Kennzahlen von Br?lhart erw?hnen?
% Erw?hnen in welcher Form die Informationen publiziert sind. Tabulation by size of income. Not on individual or household level

% http://www.estv.admin.ch/dokumentation/00075/00076/00701/index.html?lang=de
% Kantone und Gemeinden: 2010 bis 1947/1948
% Klassierte Tabellen für Abzüge, Einkommen (Steuerbares Einkommen und Reineinkommen) und Steuererträge
% Normalfälle und Sonderfälle

% Statistsiche Kennzahlen
% Mit Belastung durch die direkte Bundessteuer: 1973/1974 bis 2010
% Mit und ohne Belastung durch die direkte Bundessteuer: 1995/1996 bis 2010
% steuerbares Einkommen/Reineinkommen
% Jeweils Normal und Sonderfälle
% Individual-Daten (nicht klassierte daten)
% äquivalenzeinkommen

@@

Most studies on inequality focus on income inequality solely. However, recent activities emphasize the need of a broader conceptualization. A recent publication from the \citet{oecd_oecd_2013 UNSICHER} condense these ideas into the ICW framework (income, consumption and wealth), which is meant to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the expert group in developing the ICW Framework presented in this publication. Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_???? UNSICHER}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}}. According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being. To gain policy relevant insight, it is recommended to look at the distribution of all three distributions simultaneously. Some households with low income, for example, may report adequate levels of consumption expenditure or wealth holdings, or vice-versa. But it is also stated \citep[18]{oecd_oecd_2013 UNSICHER}:"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by collecting and providing a database following this broader concept of economic well-being. \url{http://www.lisdatacenter.org/our-data/lws-database/}} \\

@@

The assessment of income inequality is influenced by the definition of the income itself. Market income or disposable income for example differ by substantial meaning and by the expected degree of inequality. Therefore the awareness of the analyzed concept is crucial. Terminology can slightly differ, while common concepts can be identified (for detailed discussion see: \citet[44]{oecd_oecd_2013 UNSICHER}, \citet[24]{united_nations_canberra_2011}). Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution, which eventually lead to disposable income: the income measure, which finally shapes the possibility to consume. Within this framework common other income definitions are situated. These definitions are contrasted with income concepts, which can be derived out of the FTA tax data.

@@

The central income reported through tax statistics is the taxable income. It includes all reported incomes (income from employment, income from property and received transfers\footnote{Means-tested benefits are not taxed and therefore not included in tax data. Income for low income groups are therefore underestimated. However, \citep{piketty_income_2003} note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had only a trivial impact on top income shares.}) minus several deductions. It is therefore neither a pre-transfer income nor a post-transfer income measure. It's rather something in between. As the FTA tax statistics include some but not all deductions\footnote{The difference between the real total income and the taxable income are deductions. These include: professional expenses, travel expenses, interest on debt, alimonies, training costs, two-earner deduction, party contributions, private pension provision ``Säule 3a'', buying into the pension plan and sideline deductions} it is possible to calculate a sort of "total income". As some deductions can be interpreted as compulsory expenses similar to taxes the step towards total income is a step away from the income, which can be used for consumption. Similar when calculating the disposable income out of the taxable income through accounting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income). Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches, which represent the bulk of taxes in Switzerland, are missing.   \\   
%RF (16.06.2014): den letzten abschnitt finde ich etwas schwer zu verstehen als jemand der nicht tief in der materie ist. vielleicht könnte man sowas machen: It is possible to calculate a kind of "total income" that is in between net and gross income. The principle applied is to only deduct expenditures that are unavoidable.

% Die Deductions sind noch nicht gänzlich sauber abgebildet. Auf der Bilanz-Seite steht, zum steuerbaren Einkommen kommt man, indem vom Reineinkommen die Sozialabzüge(Kinder- und Fremdbetreung) sowie der Abzug für unterstützungsbedürftige Personen abzogen werden. 
% Auf den ESTV-Tabellen sind jedoch noch mehr Abzüge abgebildet. Was soll/muss/können wir nun abziehen? Bin verwirrt...

@@

The agreed standard on the statistical units, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013 UNSICHER}. Indeed it is the individuals, who receive income, own assets and experience economic well-being, but their possibility to do so, is strongly tied to the concept of household. This comprises all persons under the same housing arrangement. The basic underlying assumption for collecting data on household level instead of individual level is, that people in the same household share resources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every household member (also non-earing members, like children). Additionally, there are economies of scale when people share living space and commodities and they therefore benefit from the sharing. To compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013 UNSICHER}, \citealt{buhmann_equivalence_1988}). \\

@@

% Leigh mit Seitenzahl zitieren. Auf deutsch würde ich noch ebd.

@@

\textbf{deductions} \citet[477]{dell_income_2007}:" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
\citet[5]{schaltegger_evolution_2011 UNSICHER}: ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''. Können wir das auch? Zumindest für gewisse Zeiträume? Das wäre noch gut. \\

>>>>

% Nicht angesprochen sind in der Einleitung die Punkte 4 und 5 - Population coverage und intertemporal comparison. Der letzte Satz gefällt mir auch nicht so...

@@

% Verschiedene Punkte werden zweimal erwähnt (hier und in den folgenden Abschnitten). Man muss mal alles am Stück lesen und schauen, ob es so passt.

% Verweis auf Brülhartdaten.

@@

Most studies on inequality focus on income inequality solely. However, recent activities emphasize the need of a broader conceptualization. A recent publication from the \citet{oecd_oecd_2013} condense these ideas into the ICW framework (income, consumption and wealth), which is meant to be an internationally agreed framework on micro-level statistics.\footnote{Harmonization with other international standards was an important objective that guided the work of the expert group in developing the ICW Framework presented in this publication. Considered main standards were the System of National Accounts \citep{sna_system_2008}, the Canberra Group Handbook on Household Income Statistics \citep{united_nations_canberra_2011}, the final report of the 17th International Conference of Labour Statisticians \citep{international_labour_organisation_ilo_final_2004}} and the UNECE/CES recommendations for the 2010 Censuses of Population and Housing \citep{unece_conference_2006}}. According to the framework it is best to look at income, consumption and wealth as three separate but interrelated dimensions of people's economic well-being. To gain policy relevant insight, it is recommended to look at the distribution of all three distributions simultaneously. Some households with low income, for example, may report adequate levels of consumption expenditure or wealth holdings, or vice-versa. But it is also stated \citep[18]{oecd_oecd_2013}:"[...] integrated analysis at the household level has significant data requirements that go beyond the measurement efforts currently undertaken in most countries."\footnote{The Luxembourg Wealth Study Database is currently facing this shortcomings by collecting and providing a database following this broader concept of economic well-being. \url{http://www.lisdatacenter.org/our-data/lws-database/}} \\

@@

The assessment of income inequality is influenced by the definition of the income itself. Market income or disposable income for example differ by substantial meaning and by the expected degree of inequality. Therefore the awareness of the analyzed concept is crucial. Terminology can slightly differ, while common concepts can be identified (for detailed discussion see: \citet[44]{oecd_oecd_2013}, \citet[24]{united_nations_canberra_2011}). Figure~\ref{fig:incdef} shows a stylized framework, which includes a distinction of common income sources \footnote{Income from production of household services for own consumption is excluded because this income is hard to measure and not covered in the FTA tax data} and shows the central steps of redistribution, which eventually lead to disposable income: the income measure, which finally shapes the possibility to consume. Within this framework common other income definitions are situated. These definitions are contrasted with income concepts, which can be derived out of the FTA tax data.

@@

The central income reported through tax statistics is the taxable income. It includes all reported incomes (income from employment, income from property and received transfers\footnote{Means-tested benefits are not taxed and therefore not included in tax data. Income for low income groups are therefore underestimated. However, \citep{piketty_income_2003} note that non-taxable social security benefits grew as a share of personal income in the US but find that these changes had only a trivial impact on top income shares.}) minus several deductions. It is therefore neither a pre-transfer income nor a post-transfer income measure. It's rather something in between. As the FTA tax statistics include some but not all deductions\footnote{The difference between the real total income and the taxable income are deductions. These include: professional expenses, travel expenses, interest on debt, alimonies, training costs, two-earner deduction, party contributions, private pension provision ``Säule 3a'', buying into the pension plan and sideline deductions} it is possible to calculate a sort of total income, which is called ``net inomce'' (Reineinkommen). As some deductions can be interpreted as compulsory expenses similar to taxes the step towards total income is a step away from the income, which can be used for consumption. Similar when calculating the disposable income out of the taxable income through accounting the reported federal taxes, this is a step towards the income, which is left in the basket for consumption (disposable income). Again it is not a "pure" disposable income, because cantonal, municipal taxes and taxes from churches, which represent the bulk of taxes in Switzerland, are missing.   \\   
%RF (16.06.2014): den letzten abschnitt finde ich etwas schwer zu verstehen als jemand der nicht tief in der materie ist. vielleicht könnte man sowas machen: It is possible to calculate a kind of "total income" that is in between net and gross income. The principle applied is to only deduct expenditures that are unavoidable.
%OH (16.06.2014): bin mir nicht sicher, ob's mit gross income einfacher ist (kommt in der Grafik nicht vor)

@@

The agreed standard on the statistical units, which should be the base of inequality analysis, are households not individuals \citep[60]{oecd_oecd_2013}. Indeed it is the individuals, who receive income, own assets and experience economic well-being, but their possibility to do so, is strongly tied to the concept of household. This comprises all persons under the same housing arrangement. The basic underlying assumption for collecting data on household level instead of individual level is, that people in the same household share resources and therefore pool their incomes (when two or more earners live together) and/or use the household income to provide the essentials of living for every household member (also non-earing members, like children). Additionally, there are economies of scale when people share living space and commodities and they therefore benefit from the sharing. To compare the individual economic well-being among individuals living in different households usually equivalence scales are used (see \citealt[173]{oecd_oecd_2013}, \citealt{buhmann_equivalence_1988}). \\

@@

% Auf deutsch würde ich noch ebd.

@@

\textbf{deductions} \citet[477]{dell_income_2007}:" we can check with statistics for 1971-72 (as well as later years) presented both by size of income before deductions and income after deductions that adding back deductions does not introduce any significant error in our estimates."
\citet[5]{schaltegger_evolution_2011}: ``..., information on [...] deductions is provided in the tax statistics, thus, we could add the personal deductions to the income data to obtain a consistent series over time''. Können wir das auch? Zumindest für gewisse Zeiträume? Das wäre noch gut. \\
